Journal and article coverage

### 2024-11-28 analysis/results.Rmd
Safeguard

# Full OA journals detected through OA proportion > .95
hoad_oa_excluded <- bq_table_download("subugoe-collaborative.hoaddata.cc_oa_prop") |>
  distinct(issn_l)

@@

wos_active_jns_with_oa <- wos_active_jns |> 
  filter(oa_status == "hybrid")

@@

```{r, fig.asp=0.7}
## Combine to patchwork
layout <- "
A
B
C
"
wrap_plots(
  A = hoad_publisher_panel,
  B = first_au_publisher_panel,
  C = cor_au_publisher_panel,
  D = guide_area(),
  design = layout,
  guides = "collect",
  axes = "collect",
  heights = c(1, 2, 2)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```

@@

>>>>

# Full OA journals detected through OA proportion > .95
hoad_oa_excluded <- bq_table_download("hoa-article.hoaddata_sep24.cc_oa_prop") |>
  distinct(issn_l)

@@

wos_active_jns_with_oa <- wos_active_jns |> 
  filter(grepl("hybrid", oa_status))

@@

```{r, fig.asp=1, width = 12}
## Combine to patchwork
layout <- "
A
B
C
"
wrap_plots(
  A = hoad_publisher_panel,
  B = first_au_publisher_panel,
  C = cor_au_publisher_panel,
  D = guide_area(),
  design = layout,
  guides = "collect",
  axes = "collect",
  heights = c(1, 2, 2)
) &
  theme(
    legend.position = "top",
    legend.justification = "right",
    legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```

@@

### shared

```{r}
# Create a shared corpus via DOI
wos_dois_active_jns_with_oa_core_df <- wos_active_jns |> 
  filter(issn_l %in% wos_active_core_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  mutate(is_hybrid_wos = grepl("hybrid", oa_status)) |>
  distinct(doi, is_hybrid_wos) |>
  mutate(wos = TRUE)

scp_dois_active_jns_with_oa_core_df <- scp_active_jns |> 
  filter(issn_l %in% scp_active_core_oa$issn_l) |>
  filter(grepl("^Article|^Review", item_type)) |>
  mutate(is_hybrid_scp = grepl("hybrid", oa_status)) |>
  distinct(doi, is_hybrid_scp) |>
  mutate(scopus = TRUE)

hoad_dois_active_jns_with_oa <- cr_df |> 
  filter(issn_l %in% active_jns_with_oa$issn_l, between(cr_year, 2019, 2023)) |>
  mutate(is_hybrid_hoad = doi %in% hoaddata::cc_articles$doi) |>
  distinct(doi, issn_l, is_hybrid_hoad) |>
  mutate(hoad = TRUE)

## Shared corpus
all_dois <- dplyr::full_join(hoad_dois_active_jns_with_oa, scp_dois_active_jns_with_oa_core_df, by = "doi") |>
  dplyr::full_join(wos_dois_active_jns_with_oa_core_df, by = "doi") |>
  mutate(across(where(is.logical), ~ replace_na(.x, FALSE))) |>
  filter(!is.na(doi))

shared_dois <- all_dois |>
  filter(wos == TRUE, scopus == TRUE, hoad == TRUE)

## Merge Publisher info

shared_dois_df <- hoaddata::jct_hybrid_jns |>
  distinct(issn_l, esac_publisher) |>
  inner_join(shared_dois, by = "issn_l")

esac_five_years <- shared_dois_df |>
  group_by(esac_publisher) |>
  summarise(hoad = sum(hoad == TRUE),
            hoad_oa = sum(is_hybrid_hoad == TRUE),
            wos = sum(wos == TRUE),
            wos_oa = sum(is_hybrid_wos == TRUE),
            scp = sum(scopus == TRUE),
            scp_oa = sum(is_hybrid_scp == TRUE)) |>
  mutate(hoad_oa_prop = hoad_oa / hoad,
         wos_oa_prop = wos_oa / wos,
         scp_oa_prop = scp_oa / scp) |>
   mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           ))

oa_plot_df <- esac_five_years |> 
    pivot_longer(cols = contains(c("wos", "scp")))

oa_absolut_plot <- oa_plot_df |>
  filter(name %in% c("wos_oa", "scp_oa")) |>
  mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )
  ) |>
  mutate(publisher =
           forcats::fct_relevel(publisher, c(
             "Elsevier", "Springer Nature", "Wiley", "Other"
           ))) |>
  ggplot(aes(x = hoad_oa, y = value, size = hoad)) +
  geom_point(aes(fill = publisher), pch = 21) +
  scale_x_continuous(limits = c(0, max(
    c(
      esac_five_years$scp_oa,
      esac_five_years$hoad_oa,
      esac_five_years$wos_oa
    )
  )),
  labels = scales::number_format(big.mark = ",")) +
  scale_y_continuous(limits = c(0, max(
    c(
      esac_five_years$scp_oa,
      esac_five_years$hoad_oa,
      esac_five_years$wos_oa
    )
  )),
  labels = scales::number_format(big.mark = ",")) +
  scale_fill_manual(
    "",
    values = c(
      "Elsevier" = "#e9711c",
      "Springer Nature" = "#486a7e",
      "Wiley" = "#068853",
      "Other" = "#b3b3b3"
    )
  ) +
  facet_grid( ~ name) +
  geom_abline(slope = 1,
              intercept = 0,
              color = "blue") +
  coord_fixed() +
  theme_minimal(my_base_size)

oa_share_plot <- oa_plot_df |>
  filter(name %in% c("wos_oa", "scp_oa")) |>
   mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           )) |>
  ggplot(aes(x = hoad_oa / hoad, y = value / hoad, size = hoad)) +
  geom_point(aes(fill = publisher),  pch = 21) +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format())+
  scale_fill_manual(
      "",
      values = c(
        "Elsevier" = "#e9711c",
        "Springer Nature" = "#486a7e",
        "Wiley" = "#068853",
        "Other" = "#b3b3b3"
      )
    ) +
  facet_grid(~name) +
  geom_abline(slope = 1, intercept = 0, color = "blue") +
  coord_fixed() +
  theme_minimal(my_base_size)
```

```{r fig_oa_compare_scatter, fig.asp=1}
layout <- "
C
A
B
"
wrap_plots(
  A = oa_absolut_plot,
  B = oa_share_plot,
#  C = guide_area(),
  design = layout,
  guides = "collect",
  axes = "collect",
  heights = c(0.5, 1, 1)
) &
  theme(
   # legend.position = "top",
  #  legend.justification = "right",
 #   legend.direction = "horizontal",
 #   panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "grey50", fill = NA),
 #   axis.title=element_text(size = 10)
 ) &
    plot_annotation(tag_levels = 'A')
```

###

crossref / unpaywall

### TA countries

replication database vs database
validation shared corpus based on dois
